### 2020-04-02 Thesis-Chapter2-Background.tex
Expand on current context

Electronic encryption's dual role as enabler of strong privacy and cornerstone of computer security may be the central
conflict in the exceptional access debate. Those who focus on its role in privacy object to its absolute power. Those
who focus on its role in security rely on its absolute power.

@@

The second development was the events of and reaction to the morning of September 11, 2001. When the U.S. had already
been expanding police surveillance activities to combat communism, crime, drugs, and terrorism, the scale and audacity
of the September 11 attacks to some justified yet greater investigatory powers \cite{bloss_escalating_2007}. October of
2001 saw passage of the USA PATRIOT Act, which had several significant impacts: it weakened the limitation for using
FISA requests on U.S. citizens; it expanded the scope of what could be compelled via FISA orders; it authorized
``roving'' wiretaps; and it increased the power of National Security Letters (NSLs), which can be used without judicial
review to compel information from electronic service providers while precluding any public release of the event
\cite{sensenbrenner_2001} \cite{shamsi_2011}.

@@

The U.S. corporate response to Snowden should not be overly construed as a morally motivated defense of civil rights, as
their behavior is a matter of several practical factors \cite{treguer_us_2018}, not least of which is the pursuit of
basic digital security. Recall encryption and cryptography's critical role in security as described in
\mysec{sec-crypto-cwi}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
security is a central conflict in the exceptional access debate. When questioned by lawmakers, technical leaders and
executives have repeatedly appealed to the fact that strong cryptography is a necessity for security
\cite{schulze_clipper_2017}.

@@

\paragraph*{Apple vs. FBI} The Apple v. FBI case over access to the San Bernardino shooter's iPhone
\cite{schulze_clipper_2017}.

@@

\paragraph*{``Security vs Security''} ``Security vs Security'' vs ``Security vs Privacy''
\cite{stalla_bourdillon_privacy_2014} \cite{schneier_2019}.

>>>>

Electronic encryption's dual role as enabler of strong privacy and cornerstone of computer security may be the central
conflict in the EA debate. Those who focus on its role in privacy object to its absolute power. Those
who focus on its role in security rely on its absolute power.

@@

The second development was the events of and reaction to the morning of September 11, 2001. When the U.S. had already
been expanding police surveillance activities to combat communism, crime, drugs, and terrorism, the scale and audacity
of the September 11 attacks to some justified yet greater investigatory powers \cite{bloss_escalating_2007}. October of
2001 saw passage of the USA PATRIOT Act, which had several significant impacts: it weakened the limitation for using
FISA requests on U.S. citizens; it expanded the scope of what could be compelled via FISA orders; it authorized
``roving'' wiretaps; and it increased the power of National Security Letters (NSLs), which can be used without judicial
review to compel information from electronic service providers while precluding any public release of the event
\cite{sensenbrenner_2001} \cite{shamsi_2011}. The government was performing more surveillance than ever, and even these
weakened limitations would be repeatedly violated \cite{shamsi_2011} \cite{tucker_2020}.

@@

The U.S. corporate response to Snowden should not be overly construed as a morally motivated defense of civil rights, as
their behavior is a matter of several practical factors \cite{treguer_us_2018}, not least of which is the pursuit of
basic digital security. Recall encryption and cryptography's critical role in security as described in
\mysec{sec-crypto-cwi}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
security is a central conflict in the EA debate. When questioned by lawmakers, technical leaders and
executives have repeatedly appealed to the fact that strong cryptography is a necessity for security
\cite{schulze_clipper_2017}.

@@

\paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as antagonistic,
an assumption implicit in in the ``nothing-to-hide'' argument, but discussions in recent years have shown that the
debate is more about ``Security vs. Security'' than ``Security vs. Privacy'' \cite{stalla_bourdillon_privacy_2014}. In
this context, there are two types of security: public security, the pursuit of law enforcement, and cybersecurity, the
pursuit of tech professionals. Both sides of the debate have by now acknowledged that privacy and public security are
not always in conflict, and in fact EA's effects on public security vs. cybersecurity--which in return affects public
security--may be more important \cite{schneier_2019}.

\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack in which the suspect was killed, leaving behind a
locked iPhone, resulted in the first major battle of the Second Crypto War. The FBI issued an order under the All Writs
Act to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
objected on grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
\cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the political world's attention from
February 16, 2016, when the warrant was issued, to March 28, when the FBI announced they had gained access to the phone
through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto wars,
terrorism and device encryption. Shortly after the case ended, Senators Dianne Feinstein and Richard Burr released a
discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not make it to the
floor, and the unresolved nature of the court case left the legal debate wide open.

@@
